---
title: "CUNY Data 621 HW3 Logiistical Regression"
author: "Raphael Nash"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

#Introduction
This assignement we will do a logistical regression to predict crime giving an input dataset.   A training and evaluation dataset have been provided.   

To accomplish this I am going to use the following libraries:
1. ggplot2
2. reshape2
3. coreplot
4. forecast
5. dplyr
6. Deducer


```{r message=FALSE, warning=FALSE, echo=FALSE}

library(ggplot2)
library(reshape2)
library(corrplot) 
library(forecast)
library(dplyr)
library(Deducer)

crime_df <- read.csv("crime-training-data.csv")
```

#Data Exploration 

Summary Statistics:
```{r echo=FALSE}
summary(crime_df )
```



Box Plot of data
```{r fig.width=20,fig.height=20 , message=FALSE, warning=FALSE, echo=FALSE}

ggplot(data = melt(crime_df), aes(x=variable, y=value)) + geom_boxplot()  + coord_flip() + theme(text = element_text(size=40),
        axis.text.x = element_text(angle=70, hjust=1))
```

Correlation Plot of Data:
```{r message=FALSE, warning=FALSE, echo=FALSE }
M <- cor((crime_df)) 
corrplot(M, method = "ellipse")
```

Correlation Matrix:
```{r message=FALSE, warning=FALSE, echo=FALSE}
M
```

Scatterplats for each variable against the target:
```{r message=FALSE, warning=FALSE, echo=FALSE}
pairs(crime_df, col=crime_df$target)
```


#Transformations

Since the data does not look normally distributed, I am going to perform a Box-Cox transfomration on each of the input variables.  The labdas for each variable are as follows:
```{r message=FALSE, warning=FALSE, echo=FALSE } 
calculate_labbdas <- function(df){
  df <- df[,1:ncol(df)]
  l1 <- numeric(ncol(df))
  for (i in 1:ncol(df)){
    l1[i] <- BoxCox.lambda(df[,i])
  }
  return(data.frame(l1, colnames(df)))
}
 
box_cox_lambdas <- calculate_labbdas( dplyr::select(crime_df, -target))
colnames(box_cox_lambdas) <- c("lambda", "variable")

box_cox_lambdas
```


```{r message=FALSE, warning=FALSE, echo=FALSE}

crime_transformed_df <- crime_df
crime_transformed_df$zn <- crime_transformed_df$zn ^ (filter(box_cox_lambdas, variable=="zn")$lambda)
crime_transformed_df$indus <- crime_transformed_df$indus ^ (filter(box_cox_lambdas, variable=="indus")$lambda)
crime_transformed_df$chas <- crime_transformed_df$chas ^ (filter(box_cox_lambdas, variable=="chas")$lambda)
crime_transformed_df$nox <- crime_transformed_df$nox ^ (filter(box_cox_lambdas, variable=="nox")$lambda)
crime_transformed_df$nox <- crime_transformed_df$nox ^ (filter(box_cox_lambdas, variable=="rm")$lambda)
crime_transformed_df$age <- crime_transformed_df$age ^ (filter(box_cox_lambdas, variable=="age")$lambda)
crime_transformed_df$dis <- crime_transformed_df$dis ^ (filter(box_cox_lambdas, variable=="dis")$lambda)
crime_transformed_df$rad <- crime_transformed_df$rad ^ (filter(box_cox_lambdas, variable=="rad")$lambda)
crime_transformed_df$tax <- crime_transformed_df$tax ^ (filter(box_cox_lambdas, variable=="tax")$lambda)
crime_transformed_df$ptratio <- crime_transformed_df$ptratio ^ (filter(box_cox_lambdas, variable=="ptratio")$lambda)

```


#Build Models

I will bild a model starting with all the variables are removing the least signifigant variables until the AIC starts increaseing 


Summary of model will all varirables (model1):  
```{r message=FALSE, warning=FALSE, echo=FALSE}
model1 <- glm(target~., data =crime_transformed_df, family=binomial )
summary(model1)
```


If I drop all non signifigant variables I am left with the following variables:nox, age, dis, pratio, mdev.
Therefore I am going to build a model with thoses variables.  
Here is the summary for that model (model2)
```{r message=FALSE, warning=FALSE}
model2 <- glm(target~nox+ age+dis+ rad+ptratio+medv  , data =crime_transformed_df, family=binomial )
summary(model2)
```

The least signifigant variable of all the variables left is age, so I will drop that variable and create a model with the remaining variables.  
Here is the summary of that model:
```{r message=FALSE, warning=FALSE, echo=FALSE}
model3 <- glm(target~nox+dis+ rad+ptratio+medv , , data =crime_transformed_df, family=binomial )
summary(model3)
```
Since AIC started to go up, I am going to stop removing variables.  

#Select Models:
I am going to select the model based on area under the ROC curve (A/K/A AUC) and AIC.  

```{r message=FALSE, warning=FALSE, echo=FALSE}
rocplot(model1)
```
The AIC for model1 is `r  summary(model1)$aic`

Model2 Variables in Model 2:  nox + age + dis + rad + ptratio + medv
```{r message=FALSE, warning=FALSE, echo=FALSE }
rocplot(model2)
```

The AIC for model1 is `r  summary(model2)$aic`

Model3 Variables: nox + age + dis + rad + ptratio + medv, 
```{r message=FALSE, warning=FALSE, echo=FALSE } 
rocplot(model3)
```
The AIC for model3 is `r  summary(model3)$aic`

Based the fact that the area under the curve for model 1 and model 2 are virtually identical and the AIC for model 2 is about 1/2 the AIC for model 1 I am going to select model2.


#Make Predications
```{r message=FALSE, warning=FALSE, echo=FALSE }
crime_eval_df <- read.csv("crime-evaluation-data.csv")
crime_eval_transformed_df <- crime_eval_df
crime_eval_transformed_df$zn <- crime_eval_transformed_df$zn ^ (filter(box_cox_lambdas, variable=="zn")$lambda)
crime_eval_transformed_df$indus <- crime_eval_transformed_df$indus ^ (filter(box_cox_lambdas, variable=="indus")$lambda)
crime_eval_transformed_df$chas <- crime_eval_transformed_df$chas ^ (filter(box_cox_lambdas, variable=="chas")$lambda)
crime_eval_transformed_df$nox <- crime_eval_transformed_df$nox ^ (filter(box_cox_lambdas, variable=="nox")$lambda)
crime_eval_transformed_df$nox <- crime_eval_transformed_df$nox ^ (filter(box_cox_lambdas, variable=="rm")$lambda)
crime_eval_transformed_df$age <- crime_eval_transformed_df$age ^ (filter(box_cox_lambdas, variable=="age")$lambda)
crime_eval_transformed_df$dis <- crime_eval_transformed_df$dis ^ (filter(box_cox_lambdas, variable=="dis")$lambda)
crime_eval_transformed_df$rad <- crime_eval_transformed_df$rad ^ (filter(box_cox_lambdas, variable=="rad")$lambda)
crime_eval_transformed_df$tax <- crime_eval_transformed_df$tax ^ (filter(box_cox_lambdas, variable=="tax")$lambda)
crime_eval_transformed_df$ptratio <- crime_eval_transformed_df$ptratio ^ (filter(box_cox_lambdas, variable=="ptratio")$lambda)

probs <- predict(model2,crime_eval_df)
prediction <- ifelse ( probs > .5 ,1,0)

cbind(prediction, crime_eval_df)
```


#Appendix (R Code)
##Setup

library(ggplot2)

library(reshape2)

library(corrplot) 

library(forecast)

library(dplyr)

library(Deducer)

crime_df <- read.csv("crime-training-data.csv")


##Data Exploration 

summary(crime_df )

ggplot(data = melt(crime_df), aes(x=variable, y=value)) + geom_boxplot()  + coord_flip() + theme(text = element_text(size=40),
        axis.text.x = element_text(angle=70, hjust=1))

M <- cor((crime_df)) 
corrplot(M, method = "ellipse")

pairs(crime_df, col=crime_df$target)


##Transformations

calculate_labbdas <- function(df){

  df <- df[,1:ncol(df)]
  
  l1 <- numeric(ncol(df))
  
  for (i in 1:ncol(df)){
  
    l1[i] <- BoxCox.lambda(df[,i])
    
  }
  
  return(data.frame(l1, colnames(df)))
  
}

 
box_cox_lambdas <- calculate_labbdas( dplyr::select(crime_df, -target))

colnames(box_cox_lambdas) <- c("lambda", "variable")



box_cox_lambdas


crime_transformed_df <- crime_df

crime_transformed_df$zn <- crime_transformed_df$zn ^ (filter(box_cox_lambdas, variable=="zn")$lambda)

crime_transformed_df$indus <- crime_transformed_df$indus ^ (filter(box_cox_lambdas, variable=="indus")$lambda)

crime_transformed_df$chas <- crime_transformed_df$chas ^ (filter(box_cox_lambdas, variable=="chas")$lambda)

crime_transformed_df$nox <- crime_transformed_df$nox ^ (filter(box_cox_lambdas, variable=="nox")$lambda)

crime_transformed_df$nox <- crime_transformed_df$nox ^ (filter(box_cox_lambdas, variable=="rm")$lambda)

crime_transformed_df$age <- crime_transformed_df$age ^ (filter(box_cox_lambdas, variable=="age")$lambda)

crime_transformed_df$dis <- crime_transformed_df$dis ^ (filter(box_cox_lambdas, variable=="dis")$lambda)

crime_transformed_df$rad <- crime_transformed_df$rad ^ (filter(box_cox_lambdas, variable=="rad")$lambda)

crime_transformed_df$tax <- crime_transformed_df$tax ^ (filter(box_cox_lambdas, variable=="tax")$lambda)

crime_transformed_df$ptratio <- crime_transformed_df$ptratio ^ (filter(box_cox_lambdas, variable=="ptratio")$lambda)



##Build Models

model1 <- glm(target~., data =crime_transformed_df, family=binomial )

summary(model1)

model2 <- glm(target~nox+ age+dis+ rad+ptratio+medv  , data =crime_transformed_df, family=binomial )

summary(model2)

model3 <- glm(target~nox+dis+ rad+ptratio+medv , , data =crime_transformed_df, family=binomial )

summary(model3)

##Select Models:

rocplot(model1)

rocplot(model2)

rocplot(model3)

##Make Predications

crime_eval_df <- read.csv("crime-evaluation-data.csv")

crime_eval_transformed_df <- crime_eval_df

crime_eval_transformed_df$zn <- crime_eval_transformed_df$zn ^ (filter(box_cox_lambdas, variable=="zn")$lambda)

crime_eval_transformed_df$indus <- crime_eval_transformed_df$indus ^ (filter(box_cox_lambdas, variable=="indus")$lambda)

crime_eval_transformed_df$chas <- crime_eval_transformed_df$chas ^ (filter(box_cox_lambdas, variable=="chas")$lambda)

crime_eval_transformed_df$nox <- crime_eval_transformed_df$nox ^ (filter(box_cox_lambdas, variable=="nox")$lambda)

crime_eval_transformed_df$nox <- crime_eval_transformed_df$nox ^ (filter(box_cox_lambdas, variable=="rm")$lambda)

crime_eval_transformed_df$age <- crime_eval_transformed_df$age ^ (filter(box_cox_lambdas, variable=="age")$lambda)

crime_eval_transformed_df$dis <- crime_eval_transformed_df$dis ^ (filter(box_cox_lambdas, variable=="dis")$lambda)

crime_eval_transformed_df$rad <- crime_eval_transformed_df$rad ^ (filter(box_cox_lambdas, variable=="rad")$lambda)

crime_eval_transformed_df$tax <- crime_eval_transformed_df$tax ^ (filter(box_cox_lambdas, variable=="tax")$lambda)

crime_eval_transformed_df$ptratio <- crime_eval_transformed_df$ptratio ^ (filter(box_cox_lambdas, variable=="ptratio")$lambda)


probs <- predict(model2,crime_eval_df)

prediction <- ifelse ( probs > .5 ,1,0)

cbind(prediction, crime_eval_df)




