---
title: "CUNY DATA 621 HW 5"
output:
  pdf_document: default
  html_notebook: default
Author: Raphael Nash
---


```{r message=FALSE}
library( ggplot2)
library(reshape2)
library(corrplot)
library (MASS)
library(Deducer)
library(ROCR)
library(dplyr)
```


#Basic Data Exploration
```{r}

wine_df <- read.csv('wine-training-data.csv') 

wine_df <- wine_df[,-c(1)]
```

```{r}
summary(wine_df)
```


```{r}

wine_melt_df <- melt(wine_df[,sapply(wine_df, is.numeric)])

ggplot(wine_melt_df,aes(x = value)) + 
    facet_wrap(~variable,scales = "free") + 
    geom_histogram()
```


#Explore relationship of stars
The min value of stars is 1 therefore it seems logical that missing data in stars means zero stars. 


```{r}
wine_stars_target_df <- wine_df[,c("TARGET", "STARS")] 
wine_stars_target_df[is.na(wine_stars_target_df[,"STARS"]), "STARS"] <-0

ggplot(wine_stars_target_df, aes(x=factor(STARS), y=TARGET)) + stat_summary(fun.y="mean", geom="bar") + xlab("Average Stars")
```

```{r}
ggplot(wine_stars_target_df,aes(x = STARS)) + 
    geom_histogram()

```

#Data imputations
For this dataset we will want to treat stars as if it is a categorical variable, with all missing values treated as 0.  The rest of the missing data will be replaced with the median of the value.

```{r}
str(wine_df)
```

```{r, message=FALSE, warning=FALSE}
wine_df[is.na(wine_df[,"STARS"]), "STARS"] <-0

 
wine_df <- wine_df %>% 
   mutate_all(~ifelse(is.na(.), median(., na.rm = TRUE), .))
wine_df$STARS<- as.factor(wine_df$STARS)
```


#Transforms 
All data is relatively normal.  No transforms needed.  


#Build Models


##Possion

Build a possion model with all variables
```{r}
model_posion1 <- glm(TARGET ~  . , data=wine_df,  poisson )  
 
summary(model_posion1)
```


#Build Negaive bionomial Model 
```{r}

model_neg_binomial <- glm(TARGET ~  . , data=wine_df,  quasipoisson )  
summary(model_neg_binomial)

```

#compare models
Becasue of the fact that I transformed the STARS varaiable into a categorical variable the model is not overdispersed, therefore the possion and negative binomials are the same.  

#Generate predications


```{r}

wine_eval_df <- read.csv("wine-evaluation-data.csv")
wine_eval_df[is.na(wine_eval_df[,"STARS"]), "STARS"] <-0


wine_df <- wine_df %>% 
   mutate_all(~ifelse(is.na(.), median(., na.rm = TRUE), .))

wine_eval_df$STARS <- as.factor(wine_eval_df$STARS)
```


```{r}
str(wine_df)

```
```{r}
round(predict(model_posion1, wine_eval_df))
```

